\rule{\textwidth}{4pt}\\[0.5cm] \textbf{Discrete payout profile approximation with vanilla European options}\\[0.5cm] \rule{\textwidth}{2pt}

### 2025-02-09 latex/article.tex
numerical implementation

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{hyperref}
\usepackage{setspace}
\onehalfspacing

@@

\title{\rule{\textwidth}{4pt}\\[0.5cm]
\textbf{Payout profile approximation with vanilla European options}\\[0.5cm]
\rule{\textwidth}{2pt}}
\author{\small \textbf{Vlad Kochetov} \\
    \small Faculty of Mechanics and Mathematics, \\
    \small Taras Shevchenko National University of Kyiv \\
    \small \texttt{vladkoch@knu.ua}
}
\date{\today}

@@

\begin{abstract}
This paper explores the methodology for approximating complex payout profiles using a combination of vanilla European options. The mathematical framework is presented that allows for the decomposition of arbitrary payout structures into a series of call and put options with varying strike prices. The approach is validated through numerical examples and practical applications in financial engineering. The results demonstrate the effectiveness of this method in replicating complex payouts with high accuracy while maintaining computational efficiency.
\end{abstract}

@@

\section{Literature Review}
The concept of replicating complex payouts using simpler instruments has been extensively studied in the literature. \cite{hull2018options} provides a comprehensive overview of option pricing and replication strategies. Recent work by \cite{smith2020financial} has explored the use of machine learning techniques for payout approximation, though their focus was primarily on exotic options.

\section{Methodology}
\subsection{Mathematical Framework}
Let \( V(S, T) \) represent the value of a target payout profile at maturity \( T \) as a function of the underlying asset price \( S \). We aim to approximate \( V(S, T) \) using a combination of European call and put options:

\begin{equation}
V(S, T) \approx \sum_{i=1}^{n} \alpha_i C(S, K_i, T) + \beta_i P(S, K_i, T)
\end{equation}

where \( C(S, K_i, T) \) and \( P(S, K_i, T) \) are the prices of call and put options with strike \( K_i \), and \( \alpha_i, \beta_i \) are the corresponding weights.

\subsection{Numerical Implementation}
The approximation is achieved through the following steps:
\begin{enumerate}
    \item Discretization of the payout profile
    \item Optimization of option weights
    \item Error minimization using least squares
\end{enumerate}

\section{Results}
The proposed methodology was tested on several complex payout profiles, including digital options and barrier options. The results demonstrate that the approximation error decreases significantly as the number of vanilla options increases, with a mean absolute error of less than 1\% for most cases.

\section{Conclusion}
This paper presents a robust method for approximating complex payout profiles using vanilla European options. The approach offers significant advantages in terms of computational efficiency and practical implementation, making it a valuable tool for financial engineers and practitioners.

>>>>

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{hyperref}
\usepackage{setspace}
\onehalfspacing
\DeclareMathOperator*{\argmin}{arg\,min}

@@

\title{\rule{\textwidth}{4pt}\\[0.5cm]
\textbf{Discrete payout profile approximation with vanilla European options}\\[0.5cm]
\rule{\textwidth}{2pt}}
\author{\small \textbf{Vlad Kochetov} \\
    \small Faculty of Mechanics and Mathematics, \\
    \small Taras Shevchenko National University of Kyiv \\
    \small \texttt{vladkoch@knu.ua}
}
\date{\today}

@@

\begin{abstract}
This paper explores the methodology for approximating complex payout profiles using a
combination of vanilla European options. The mathematical framework is presented that
allows for the decomposition of arbitrary payout structures into a series of call and
put options with varying strike prices. The approach is validated through numerical
examples and practical applications in financial engineering. The results demonstrate
the effectiveness of this method in replicating complex payouts.
\end{abstract}

\newpage

@@

\section{Literature Review}
Foundational work by Black and Scholes \cite{black1973pricing} established the theoretical framework for option pricing under geometric Brownian motion assumptions. Subsequent research has extended these principles to various derivative pricing applications.

\section{Methodology}
\subsection{Model Assumptions}
The analysis assumes the underlying asset follows geometric Brownian motion with constant drift $\mu$ and volatility $\sigma$ \cite{black1973pricing}:
\begin{equation}
dS_t = \mu S_t dt + \sigma S_t dW_t
\end{equation}
where $W_t$ is a Wiener process. Option pricing utilizes the Black-Scholes framework \cite{black1973pricing} for fair value calculation, incorporating risk-neutral valuation principles.

\subsection{Mathematical Framework}
Let \( V(S, T) \) represent the value of a target payout profile at maturity \( T \)
as a function of the underlying asset price \( S \). We construct an approximation
using \( n \) call options and \( m \) put options with strike prices \( \{K_i\}_{i=1}^{n+m} \):

\begin{equation}
V(S, T) \approx \lambda S + \sum_{i=1}^{n} \alpha_i C(S, K_i, T) + \sum_{j=1}^{m} \beta_j P(S, K_j, T)
\end{equation}

where \( \lambda \) represents the spot position, \( \alpha_i \) and \( \beta_j \) are
option weights. The optimal weights are found by solving the regularized least squares
problem:

\begin{equation}
\min_{\boldsymbol{\theta}} \int_{S_{\min}}^{S_{\max}} \left[ V(S) - \hat{V}(S;\boldsymbol{\theta}) \right]^2 dS + \gamma \|\boldsymbol{\theta}\|_2^2
\end{equation}

where \( \boldsymbol{\theta} = (\lambda, \{\alpha_i\}, \{\beta_j\}) \) represents
the position weights, \( \gamma \) is the regularization parameter, and
\( S_{\min}, S_{\max} \) define the approximation domain.

Discretizing the integral and using matrix notation, we solve:

\begin{equation}
\boldsymbol{\theta}^* = \argmin_{\boldsymbol{\theta}} \|\mathbf{A}\boldsymbol{\theta} - \mathbf{b}\|_2^2 + \gamma \|\boldsymbol{\theta}\|_2^2
\end{equation}

where the design matrix \( \mathbf{A} \) contains option payoffs and spot positions
evaluated at discrete price points \( \{S_k\}_{k=1}^N \), and \( \mathbf{b} \) is
the vector of target payout values.

\subsection{Numerical Implementation}
The discrete optimization problem is solved using regularized least squares:
\begin{equation}
\boldsymbol{\theta}^* = (\mathbf{A}^\top \mathbf{A} + \gamma \mathbf{I})^{-1} \mathbf{A}^\top \mathbf{b}
\end{equation}

where:
\begin{itemize}
\item \( \mathbf{A} \in \mathbb{R}^{N \times (n+m+1)} \): Matrix of basis payoffs
\item \( \mathbf{b} \in \mathbb{R}^N \): Target payoff values
\item \( \gamma \): Regularization parameter controlling solution smoothness
\item \( \mathbf{I} \): Identity matrix for L2 regularization
\end{itemize}

The evaluation points \( \{S_k\} \) are typically distributed around the current
spot price to emphasize approximation accuracy in relevant price regions.

\section{Results}
\begin{figure}[htbp]
\centering
\input{regularization_comparison.tex}
\caption{Comparison of L1 and L2 regularization methods for payout profile approximation}
\label{fig:regularization}
\end{figure}

\section{Conclusion}
